---
title: "Atlas Labs HR Analytics Report"
subtitle: "Employee Attrition & Performance Analysis"
author: "Atlas Labs HR Analytics Dashboard"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
    fig_width: 10
    fig_height: 6
  pdf_document:
    toc: true
    number_sections: true
    fig_width: 10
    fig_height: 6
params:
  data_summary: NULL
  selected_filters: NULL
  analysis_date: NULL
  kpi_metrics: NULL
  attrition_analysis: NULL
  performance_data: NULL
  satisfaction_metrics: NULL
  compensation_analysis: NULL
  recommendations: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 300
)

# Load required libraries
library(tidyverse)
library(plotly)
library(DT)
library(knitr)
library(kableExtra)
```

```{css, echo=FALSE}
/* Custom CSS for professional report styling */
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}

.kpi-box {
  background: linear-gradient(135deg, #667eea 0%, #764ba2 100%);
  color: white;
  padding: 20px;
  border-radius: 10px;
  text-align: center;
  margin: 10px;
}

.recommendation-box {
  background-color: #f8f9fa;
  border-left: 4px solid #28a745;
  padding: 15px;
  margin: 10px 0;
}

h1, h2 { color: #2c3e50; }
h3, h4 { color: #34495e; }
```

# Executive Summary {.tabset}

## Key Findings

```{r executive-summary}
# Display key metrics from params
if (!is.null(params$kpi_metrics)) {
  kpi_data <- params$kpi_metrics
  
  cat("### Atlas Labs Workforce Overview\n\n")
  cat("**Analysis Period:** ", format(Sys.Date(), "%B %Y"), "\n\n")
  
  # Create KPI summary
  if ("total_employees" %in% names(kpi_data)) {
    cat("- **Total Employees:** ", kpi_data$total_employees, "\n")
  }
  if ("attrition_rate" %in% names(kpi_data)) {
    cat("- **Overall Attrition Rate:** ", 
        scales::percent(kpi_data$attrition_rate, accuracy = 0.1), "\n")
  }
  if ("avg_satisfaction" %in% names(kpi_data)) {
    cat("- **Average Job Satisfaction:** ", 
        round(kpi_data$avg_satisfaction, 1), "/5.0\n")
  }
  if ("avg_salary" %in% names(kpi_data)) {
    cat("- **Average Salary:** $", 
        scales::comma(kpi_data$avg_salary), "\n")
  }
}
```

## Strategic Recommendations

```{r recommendations}
if (!is.null(params$recommendations)) {
  for (i in seq_along(params$recommendations)) {
    cat('<div class="recommendation-box">\n')
    cat("**Recommendation ", i, ":** ", params$recommendations[[i]]$title, "\n\n")
    cat(params$recommendations[[i]]$description, "\n\n")
    cat("*Priority:* ", params$recommendations[[i]]$priority, "\n")
    cat('</div>\n\n')
  }
} else {
  cat("Recommendations will be generated based on analysis results.")
}
```

# Workforce Demographics

## Employee Distribution

```{r demographics-overview}
# Placeholder for demographic analysis
if (!is.null(params$data_summary)) {
  cat("### Current Workforce Composition\n\n")
  # Add demographic charts and tables here
  cat("*Demographic analysis charts will be inserted here from the Shiny app data.*\n\n")
}
```

```{r demographics-charts, fig.cap="Employee Demographics Overview"}
# Placeholder for demographic visualizations
# These will be passed from the Shiny app modules
plot.new()
text(0.5, 0.5, "Demographics Visualization\n(Generated from Shiny Module)", 
     cex = 1.5, adj = 0.5)
```

## Diversity Metrics

```{r diversity-analysis}
cat("### Diversity & Inclusion Metrics\n\n")
cat("- Gender distribution across departments\n")
cat("- Ethnic diversity representation\n") 
cat("- Age distribution analysis\n")
cat("- Geographic spread of workforce\n\n")
```

# Attrition Analysis

## Attrition Trends

```{r attrition-overview}
if (!is.null(params$attrition_analysis)) {
  cat("### Attrition Rate Analysis\n\n")
  
  # Display attrition metrics
  attrition_data <- params$attrition_analysis
  
  if ("by_department" %in% names(attrition_data)) {
    cat("**Attrition by Department:**\n\n")
    dept_table <- attrition_data$by_department
    kable(dept_table, 
          col.names = c("Department", "Attrition Rate", "Employee Count"),
          format = "html") %>%
      kable_styling(bootstrap_options = c("striped", "hover"))
  }
}
```

```{r attrition-visualization, fig.cap="Attrition Analysis by Key Factors"}
# Placeholder for attrition visualizations
plot.new()
text(0.5, 0.5, "Attrition Analysis Visualization\n(Generated from Shiny Module)", 
     cex = 1.5, adj = 0.5)
```

## Risk Factors

```{r risk-factors}
cat("### Key Attrition Risk Factors\n\n")
cat("1. **Tenure & Role Duration**\n")
cat("   - Analysis of years in current role vs. attrition\n")
cat("   - Time since last promotion impact\n\n")

cat("2. **Compensation & Benefits**\n")
cat("   - Salary competitiveness\n")
cat("   - Stock option levels\n\n")

cat("3. **Work Environment**\n")
cat("   - Overtime requirements\n")
cat("   - Distance from home\n")
cat("   - Business travel frequency\n\n")
```

# Performance Analytics

## Performance Distribution

```{r performance-overview}
if (!is.null(params$performance_data)) {
  cat("### Performance Rating Analysis\n\n")
  cat("Analysis of employee performance ratings across various dimensions.\n\n")
}
```

```{r performance-charts, fig.cap="Performance Metrics Overview"}
# Placeholder for performance visualizations
plot.new()
text(0.5, 0.5, "Performance Analysis Visualization\n(Generated from Shiny Module)", 
     cex = 1.5, adj = 0.5)
```

## Training & Development

```{r training-analysis}
cat("### Training Opportunities Analysis\n\n")
cat("- Training opportunities offered vs. taken\n")
cat("- Impact of training on performance ratings\n")
cat("- Training gaps by department and role\n")
cat("- ROI analysis of training programs\n\n")
```

# Employee Satisfaction

## Satisfaction Metrics

```{r satisfaction-overview}
if (!is.null(params$satisfaction_metrics)) {
  cat("### Employee Satisfaction Analysis\n\n")
  
  satisfaction_data <- params$satisfaction_metrics
  
  cat("**Key Satisfaction Areas:**\n\n")
  cat("- Job Satisfaction\n")
  cat("- Environment Satisfaction\n") 
  cat("- Relationship Satisfaction\n")
  cat("- Work-Life Balance\n\n")
}
```

```{r satisfaction-visualization, fig.cap="Employee Satisfaction Metrics"}
# Placeholder for satisfaction visualizations
plot.new()
text(0.5, 0.5, "Satisfaction Analysis Visualization\n(Generated from Shiny Module)", 
     cex = 1.5, adj = 0.5)
```

## Correlation Analysis

```{r satisfaction-correlation}
cat("### Satisfaction Impact Analysis\n\n")
cat("Analysis of how satisfaction metrics correlate with:\n\n")
cat("- Employee retention\n")
cat("- Performance ratings\n")
cat("- Career progression\n")
cat("- Compensation levels\n\n")
```

# Compensation Analysis

## Salary Distribution

```{r compensation-overview}
if (!is.null(params$compensation_analysis)) {
  cat("### Compensation Analysis Overview\n\n")
  cat("Comprehensive analysis of salary distribution and equity.\n\n")
}
```

```{r compensation-charts, fig.cap="Compensation Analysis"}
# Placeholder for compensation visualizations
plot.new()
text(0.5, 0.5, "Compensation Analysis Visualization\n(Generated from Shiny Module)", 
     cex = 1.5, adj = 0.5)
```

## Pay Equity

```{r pay-equity}
cat("### Pay Equity Analysis\n\n")
cat("- Gender pay gap analysis\n")
cat("- Ethnic pay equity review\n")
cat("- Department salary benchmarking\n")
cat("- Role-based compensation comparison\n\n")
```

# Data Quality & Methodology

## Data Sources

```{r data-methodology}
cat("### Data Sources & Quality\n\n")
cat("**Primary Data Sources:**\n\n")
cat("- Employee Master Data (employee.csv)\n")
cat("- Performance Ratings (performance_rating.csv)\n")
cat("- Education Levels (education_level.csv)\n\n")

cat("**Data Quality Metrics:**\n\n")
if (!is.null(params$data_summary)) {
  if ("data_quality" %in% names(params$data_summary)) {
    quality_metrics <- params$data_summary$data_quality
    for (metric in names(quality_metrics)) {
      cat("- ", metric, ": ", quality_metrics[[metric]], "\n")
    }
  }
}

cat("\n**Analysis Period:** ", params$analysis_date %||% "Current", "\n\n")
```

## Filters Applied

```{r applied-filters}
if (!is.null(params$selected_filters) && length(params$selected_filters) > 0) {
  cat("### Active Filters\n\n")
  
  for (filter_name in names(params$selected_filters)) {
    filter_values <- params$selected_filters[[filter_name]]
    if (length(filter_values) > 0) {
      cat("- **", filter_name, ":** ", paste(filter_values, collapse = ", "), "\n")
    }
  }
} else {
  cat("### Filters Applied\n\n")
  cat("No filters applied - analysis includes all employee records.\n\n")
}
```

# Appendix

## Technical Notes

```{r technical-notes}
cat("### Technical Implementation\n\n")
cat("- **Analysis Tool:** Atlas Labs HR Analytics Dashboard\n")
cat("- **Report Generated:** ", format(Sys.time(), "%Y-%m-%d %H:%M:%S"), "\n")
cat("- **R Version:** ", R.version.string, "\n")
cat("- **Dashboard Version:** 1.0.0\n\n")

cat("### Data Definitions\n\n")
cat("- **Attrition Rate:** Percentage of employees who left the organization\n")
cat("- **Satisfaction Ratings:** Scale of 1-5 (1=Very Dissatisfied, 5=Very Satisfied)\n")
cat("- **Performance Ratings:** Scale of 1-5 (1=Poor, 5=Excellent)\n\n")
```

## Contact Information

```{r contact-info}
cat("### Report Credits\n\n")
cat("**Developed by:** akhapwoyaco\n")
cat("**GitHub:** https://github.com/akhapwoyaco\n")
cat("**Data Source:** [HR Analytics Dataset](https://herdataproject.gumroad.com/l/hr-analytics-tableau)\n\n")
cat("For questions about this analysis or the dashboard, please contact the HR Analytics team.\n\n")
```

---

*This report was automatically generated by the Atlas Labs HR Analytics Dashboard. All data has been analyzed in accordance with company privacy policies and data governance standards.*